---
title: "General social society"
author: "Andrew Arredondo"
date: "sep 10/2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->



The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Question : is gun ownership independent of opinion on capital punishment?

$H_0$: gun ownership and opinion on capital punishment are not related.
$H_A$: gun ownership and opinion on capital punishment related.

OWNGUN ~ CAPPUN both are categorical, thus is use cat~cat


## Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~OWNGUN + CAPPUN,data=GSSdata)
barchartGC(~OWNGUN + CAPPUN,data=GSSdata, type="percent")

```
This barchart shows gun owners are more likely to favor captial punishment.
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~OWNGUN + CAPPUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

75.68% gun owners favor capital punishment. 24.32% gun owners oppose Capital punishment
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The inferential results are showing me that the p-value is 0.0202
meaning it is a alternative hypothesis because the true odds ratio is not equal to 1.

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
plot(EDUC~VOTE12,data=GSSdata)
```

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Voted12 median is higher than no in 12. The education is also more in voted 12 because it shows the people who voted have more and higher education than no in 12.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```
The average education for who voted is 15.7 while the average for the people who didn't vote was 13.4 the median difference between the two is 3.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```

The inferential results show that it is a alternative hypothesis because it has a true difference in means between group no in 12 is not equal to 0. There is a relationship because there is a difference greater than 0.


#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

